Automated Program Verification and Synthesis of Embedded \& Cyber-physical Systems: Advances and Challenges

### 2017-08-03 paper.tex
reviewed section 3
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

As an example, consider a simple C program (slightly modified from~\cite{Strichman08}) with an exponential number of paths, as shown in Figure~\ref{figure:verification-condition}(a); the corresponding C program, in single static assignment (SSA) form~\cite{Appel97}, is shown in Figure~\ref{figure:verification-condition}(b). One may notice that the SSA form is an intermediate representation, which is used by compilers to facilitate optimizations and transformations of source code. The common property in SSA form is that every variable has only one definition in the program text. This is achieved by introducing a fresh variable from the original name ({\it e.g.}, with a subscript), such that every assignment has a unique left hand side, as shown in Figure~\ref{figure:verification-condition}(b).
%
\begin{figure}[ht]
\centering
\begin{minipage}{0.45\textwidth}
\begin{lstlisting}
#include <assert.h>
int x[N], a;
...
int main(void) {
  a=N;
  for(int i=0; i<N; i++)
    if (x[i]>1)
      a--;
  assert(a<=N);
  return 0;
}
\end{lstlisting}
\end{minipage}
\begin{center}
(a)
\end{center}
\centering
\begin{minipage}{0.45\textwidth}
\begin{lstlisting}
a1 = N
a2 = (x[0] > 1) ? a1 - 1 : a1
a3 = (x[1] > 1) ? a2 - 1 : a2
a4 = (x[2] > 1) ? a3 - 1 : a3
...
an+1 = (x[N-1] > 1) ? an - 1 : an 
\end{lstlisting}
\end{minipage}
\begin{center}
(b)
\end{center}
\caption{(a) A simple C program with a loop \textit{for}. (b) The corresponding unwound C program of (a) converted into SSA form.} 
\label{figure:verification-condition}
\end{figure}

@@

One promising approach to achieve completeness, in BMC techniques, is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}. The main challenge, however, of such an approach relies on computing and strengthening inductive invariants from programs. In particular, loop invariants, which are computed from programs under verification, must be inductive (and not just invariant), in order to check the corresponding verification conditions, {\it i.e.}, induction cannot determine the invariance of a non-inductive assertion~\cite{Bradley07}. 
%
For instance, consider the C-code fragment shown in Figure~\ref{figure:inductive-invariant}. If we consider the IEEE floating-point standard (IEEE 754)~\cite{4610935,Goldberg91whatevery}, then  the loop invariant $x>0$ holds initially and after each iteration; actually, $x$ tends to infinity after $128$ iterations. This invariant, however, is not inductive, given that $x>0$ before an initial iteration does not ensure that $x>0$ after each iteration; if we initially have $x=0.9$, then $x<0$ after the fourth iteration. As a consequence, even if \textit{k}-induction procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that verifiers can automatically prove program correctness. In this specific example, an inductive invariant would be $x>1$, given that if we can check that if $x>1$ before the initial iteration, then also $x>1$ after \textit{k} iterations.
%
\begin{figure}[h]
\centering
\begin{minipage}{0.45\textwidth}
\begin{lstlisting}
#include <assert.h>
int main(void) {
 float x=2;
 while (1) {
   x = ((2*x) - 1);
 }
 return 0;
}
\end{lstlisting}
\end{minipage}
\caption{Motivating example for inductive invariants.}
\label{figure:inductive-invariant}
\end{figure}

@@

Nonetheless, there is~little evidence, in the available literature, that model checking hardware and software systems through \textit{k}-induction (and invariants) can be efficiently exploited in embedded-system verification. That happens due to the distinctive characteristics mentioned earlier, which influence embedded system development and also verification processes. Additionally, there is still a lack of studies for embedded software verifiers to exploit the combination and configuration of different invariant generation and strengthening algorithms, including analysis to discover linear inequalities, polynomial equalities and inequalities, and invariants related to memory and variable aliasing~\cite{Bradley07}.

>>>>

As an example, consider a simple C program (slightly modified from~\cite{Strichman08}) with an exponential number of paths, as shown in Figure~\ref{figure:verification-condition}(a); the corresponding C program, in single static assignment (SSA) form~\cite{Appel97}, is shown in Figure~\ref{figure:verification-condition}(b). One may notice that the SSA form is an intermediate representation, which is used by compilers to facilitate optimizations and transformations of source code. The common property in SSA form is that every variable has only one definition in the program text. This is achieved by introducing a fresh variable from the original name ({\it e.g.}, with a subscript), such that every assignment has a unique left-hand side, as shown in Figure~\ref{figure:verification-condition}(b).
%
\begin{figure}[ht]
\centering
\begin{minipage}{0.45\textwidth}
\begin{lstlisting}
#include <assert.h>
int x[N], a;
...
int main(void) {
  a=N;
  for(int i=0; i<N; i++)
    if (x[i]>1)
      a--;
  assert(a<=N);
  return 0;
}
\end{lstlisting}
\end{minipage}
\begin{center}
(a)
\end{center}
\centering
\begin{minipage}{0.45\textwidth}
\begin{lstlisting}
a1 = N
a2 = (x[0] > 1) ? a1 - 1 : a1
a3 = (x[1] > 1) ? a2 - 1 : a2
a4 = (x[2] > 1) ? a3 - 1 : a3
...
an+1 = (x[N-1] > 1) ? an - 1 : an 
\end{lstlisting}
\end{minipage}
\begin{center}
(b)
\end{center}
\caption{(a) A simple C program with a loop \textit{for}. (b) The corresponding unwound C program of (a) converted into SSA form.} 
\label{figure:verification-condition}
\end{figure}

@@

One promising approach to achieve completeness, in BMC techniques, is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}. The main challenge, however, of such an approach relies on computing and strengthening inductive invariants from programs. In particular, loop invariants, which are computed from programs under verification, must be inductive (and not just invariant), in order to check the corresponding VCs, {\it i.e.}, induction cannot determine the invariance of a non-inductive assertion~\cite{Bradley07}.

For instance, consider the C-code fragment shown in Figure~\ref{figure:inductive-invariant}. If we consider the IEEE floating-point standard (IEEE 754)~\cite{4610935,Goldberg91whatevery}, then  the loop invariant $x>0$ holds initially and after each iteration; actually, $x$ tends to infinity after $128$ iterations. This invariant, however, is not inductive, given that $x>0$ before an initial iteration does not ensure that $x>0$ after each iteration; if we initially have $x=0.9$, then $x<0$ after the fourth iteration. As a consequence, even if invariant generation procedures successfully compute such assertions, which are indeed invariant, those must be inductive, so that\textit{k}-induction verifiers can automatically prove program correctness. In this specific example, an inductive invariant would be $x>1$, given that if we can check that if $x>1$ before the initial iteration, then also $x>1$ after \textit{k} iterations.
%
\begin{figure}[h]
\centering
\begin{minipage}{0.45\textwidth}
\begin{lstlisting}
#include <assert.h>
int main(void) {
 float x=2;
 while (1) {
   x = ((2*x) - 1);
 }
 return 0;
}
\end{lstlisting}
\end{minipage}
\caption{Motivating example for inductive invariants.}
\label{figure:inductive-invariant}
\end{figure}

@@

Nonetheless, there is~little evidence, in the available literature, that model checking hardware and software systems through \textit{k}-induction (and invariants) can be efficiently exploited in ECPS verification. That happens due to the distinctive characteristics mentioned earlier, which influence ECPS development and also verification processes. Additionally, there is still a lack of studies for embedded software verifiers to exploit the combination and configuration of different invariant generation and strengthening algorithms, including analysis to discover linear inequalities, polynomial equalities and inequalities, and invariants related to memory and variable aliasing~\cite{Bradley07}.
